---
title: "week 2 quiz solution"
author: 
output: pdf_document
---

Week 2 KNN review quiz:

* Use the iris dataset to create train and test sets (80% of data as train)
* Find the best k for KNN by trying k from 1 through 20 on the train set
* Plot k vs accuracy
* Make a confusion matrix of the results on the test set
* Write at least 1 sentence explaining/interpreting the confusion matrix

```{r check_out_data}
# maybe you want to do something here for reproducibility?  skip if you don't know it right away
str(iris)
```

Create a train/test split with 80% of the data going to training.
```{r train_test_split}
# create the train indices with createDataPartition
library(caret)
train.indices <- createDataPartition(iris$Species, p=0.8)$Resample1
# convert the iris dataset to a data table so we can easily create train/test
library(data.table)
set.seed(23)
iris.dt <- as.data.table(iris)
train.indices <- createDataPartition(iris.dt, p=0.8, list=FALSE)

# select the train and test sets here --  getting an error?
# print out your train.indices variable to see what you're missing
train <- iris.dt[train.indices,]
test <- iris.dt[-train.indices,]


dim(train); dim(test)

```

Fit the model here.  We use caret with repeatedcv to minimize and average out any sort of noise we might encounter due to the train/validation splits.
```{r fit_model}
# use cv or repeatedcv
trControl <- trainControl(method ="repeatedcv", number=3, repeats=50 )
set.seed(23)
?getModelInfo
fit <- train(Species ~ .,
             method     = "knn" ,
             trControl  =,
             preProcess = c("center", "scale"),
             metric     = "Accuracy",
             data       = train)


print(fit)
```


```{r accuracy_plot}
# plot accuracy vs k here
plot()
```

```{r conf_matrix}
# predict on the test set and make a confusion matrix
# don't forget to write a sentence explaining/interpreting the confusion matrix
preds <- predict()
confusionMatrix()
```