---
title: "week 2 Homework"
author: "Jeanette Henry"
output: word_document
---

Week 2 
Use the cleaned heart disease data set -- either use the data you cleaned, or the heart.disease.data.clean.csv file under week 2 (topic 2) content.

Our goal is to predict if someone has heart disease based on the features we have.

Here is the simplest way to do this assignment:

The 'num' column is our target variable we are trying to predict.   0 means little or no heart disease (constriction of blood vessels due to plaque buildup), 1-4 is increasing amounts of blood vessel constriction (4 means the worst case of heart disease).

For the num column, leave the 0s as 0, and make anything that is 1 through 4 equal to 1.  You can filter the rows of a data.table like this:
  heart.dt[num >=1, ]
to get the rows where num is greater than or equal to 1.  Then set the value of 'num' in those rows to 1.  I recommend using the := operator to set the num row by reference.
Split up the data into train/test datasets, I recommend using the createDataPartition() function from the caret library.
Tune the k hyperparameter to the optimal value and support it with data (elbow plot).  Or at least try 3 values of k and compare the performance on the test set.
Report accuracy and/or other scoring metrics (confusion matrix, R^2, SSE, etc) on train and test data sets.
Make at least 2 plots. 

```{r check_out_data}
library(caret)
library(data.table)
library(ggplot2)
set.seed(223) 
filename <- 'C:/MSDS/MachL/Wk2/heart.disease.data.clean.csv'
heart.dt <- fread(filename)

str(heart.dt)
col.labels <- c('age',
                'sex',
                'chest.pain.type',
                'resting.blood.pressure',
                'cholesterol',
                'num.cigs.per.day',
                'years.as.smoker',
                'fasting.blood.sugar',
                'family.history.heart.disease',
                'resting.ecg.results',
                'max.heart.rate.exercise',
                'exercise.induced.angina',
                'blood.disorder',  # Thalassemia I think
                'heart.disease')

names(heart.dt) <- col.labels
heart.dtog <- heart.dt
```


``` {r convert}
heart.dt[heart.disease > 0, heart.disease:=1]
heart.dt[, heart.disease:=as.factor(heart.disease)]
levels(heart.dt$heart.disease) <- c('no', 'yes')
summary(heart.dt)
dim(heart.dt)
# 13 features -- we will need this for a hyperparameter
```

Create a train/test split with 80% of the data going to training.
```{r train_test_split}
# create the train indices with createDataPartition
tr.idxs <- createDataPartition(heart.dt$heart.disease, p = 0.8)$Resample1
train <- heart.dt[tr.idxs]
test <- heart.dt[-tr.idxs]
```

Fit the model here.  We use caret with repeatedcv to minimize and average out any sort of noise we might encounter due to the train/validation splits.
```{r fit_model}
trControl <- trainControl(method  = "repeatedcv",
                          number  = 3,
                          repeats = 10,
                          classProbs=TRUE)
                          #summaryFunction = twoClassSummary)
set.seed(223)
fit <- train(heart.disease ~ .,
             data=train,
             method= 'knn',
             tuneGrid = expand.grid(k = 1:20),
             trControl= trControl,
             preProcess= c("center", "scale"),
             metric= "Accuracy")


print(fit)
plot(fit)
```



The accuracy plot ends up with around 12 being optimal for k. giving an accuracy of .791


```{r conf_matrix}
# 12 for k seems to be best
preds <- predict(fit, test)
confusionMatrix(preds, test$heart.disease, positive = 'yes')
```
Test accuracy of 0.8393 with only 9 misclassified, much better than the no information rate of 0.5536

```{r varImp}
filename <- 'C:/MSDS/MachL/Wk2/heart.disease.data.clean.csv'
heart.dtog <- fread(filename)

#str(heart.dtog)
col.labels <- c('age',
                'sex',
                'chest.pain.type',
                'resting.blood.pressure',
                'cholesterol',
                'num.cigs.per.day',
                'years.as.smoker',
                'fasting.blood.sugar',
                'family.history.heart.disease',
                'resting.ecg.results',
                'max.heart.rate.exercise',
                'exercise.induced.angina',
                'blood.disorder',  # Thalassemia I think
                'heart.disease')

names(heart.dtog) <- col.labels
str(heart.dtog)

varImp(fit, scale = F)
#blood.disorder,chest.pain.type, and max.hear.rate.exercise are big factors
str(heart.dt)
#featplot <- ggplot(heart.dtog, aes(x=blood.disorder, y=max.heart.rate.exercise, color=heart.disease)) + 
#  geom_point() + 
#  scale_color_gradient(low="blue", high="red")
#featplot
#too much overlap, add jitter to points
featplot <- ggplot(heart.dtog, aes(x=chest.pain.type, 
                                    y=max.heart.rate.exercise,
                                    color=heart.disease)) + 
  geom_jitter() + 
  scale_color_gradient(low="blue", high="red")
featplot
library(plotly)
heart.dtog[, blood.disorder:=as.numeric(blood.disorder)]
heart.dtog[, blood.disorder:=jitter(blood.disorder)]
heart.dtog[, chest.pain.type:=as.numeric(chest.pain.type)]
heart.dtog[, chest.pain.type:=jitter(chest.pain.type)]
str(heart.dtog)
featplot2 <- plot_ly(heart.dtog, 
                    x=~chest.pain.type, 
                    y=~blood.disorder,
                    z=~max.heart.rate.exercise, 
                    marker = list
                    (
                    color=~heart.disease,
                    colorscale =c('#FFE1A1', '#683531'),
                    showscale=TRUE
                    ))

```


