---
title: "week 2 Homework"
author: "Nate George"
output: pdf_document
---

Week 2 KNN quiz:

* Use the iris dataset to create train and test sets (80% of data as train)
* Find the best k for KNN by trying k from 1 through 20 on the train set
* Plot k vs accuracy
* Make a confusion matrix of the results on the test set
* Write at least 1 sentence explaining/interpreting the confusion matrix

```{r check_out_data}
set.seed(223) 
str(iris)
```

Create a train/test split with 80% of the data going to training.
```{r train_test_split}
# create the train indices with createDataPartition
library(caret)
train.indices <- createDataPartition(iris$Species, p = 0.8)$Resample1
library(data.table)
iris.dt <- as.data.table(iris)
train <- iris.dt[train.indices, ]
# this doesn't work with data.frames, only data.tables
test <- iris.dt[-train.indices]
```

Fit the model here.  We use caret with repeatedcv to minimize and average out any sort of noise we might encounter due to the train/validation splits.
```{r fit_model}
trControl <- trainControl(method  = "repeatedcv",
                          number  = 3,
                          repeats = 30)

fit <- train(Species ~ .,
             method     = 'knn',
             tuneGrid   = expand.grid(k = 1:20),
             trControl  = trControl,
             preProcess = c("center", "scale"),
             metric     = "Accuracy",
             data       = train)


print(fit)
```

The accuracy plot ends up with around 7-9 being optimal for k.  This changes depending on your random seed.
```{r accuracy_plot}
plot(fit$results$k, fit$results$Accuracy)
```

```{r conf_matrix}
# 7 or 8 for k seems to be best
preds <- predict(fit, test)
confusionMatrix(preds, test$Species)
```

It looks like all classes were perfectly classified except versicolor.  This one had one misclassification as virginica.