---
title: "week1_R_data_prep_demo"
author: "Nate George"
date: "January 9, 2018"
output: html_document
---

We are going to examine the auto mpg dataset from here:

as an example for how to prepare data.

First we'll load the libraries we will need (message=F will hide messages from the package loading):

```{r load_libs, message=F}
library(data.table)
library(DMwR)
library(corrplot)
```

message=F so that we don't get messages about loading different libraries in the PDF
echo=F if you want to hide an entrie code block
Next we'll load the data and check it out:

```{r load_data}
fn <- 'C:/MSDS/MachL/Wk1/auto-mpg.data'

# as.is leaves characters as characters instead of converting to factors, so we can substitute NA for ?
df <- read.table(fn, as.is = T)
auto.dt <- as.data.table(df)
str(auto.dt)
```

V4 is the horsepower column, and is a string because there are some missing values represented as '?'.
First we're going to replace the V-names with more accurate and easy-to-use names:

```{r set_names}
fn <- 'C:/MSDS/MachL/Wk1/auto-mpg.names'

auto.names <- readLines(fn)
auto.names

# gets the name from the list of names using regular expressions
get_name <- function(x) gsub('\\s+\\d+\\.\\s+(.+):\\s+.+', '\\1', x)
short.names <- unlist(lapply(auto.names, FUN = get_name))
short.names <- unlist(lapply(short.names, FUN = function(x) gsub('\\s', '.', x)))
short.names

names(auto.dt) <- short.names
```

Now we want to replace missing values with NA.  We can see horsepower is the only colunm with missing values, and there are 6 of them.

```{r repl_missing}
auto.dt[auto.dt == '?'] <- NA
str(auto.dt)
```

There are a ton of car names, so we will drop that column, and we need to convert the horespower to numeric.

```{r set_cols}
auto.dt[, horsepower:=as.numeric(horsepower)]
auto.dt[, car.name:=NULL]
str(auto.dt)
```

Now we will replace missing values by imputing them with K-nearest neighbors.
```{r imputation}
auto.dt.nona <- knnImputation(auto.dt)
# if you're curious, see what the nas were replaced with
auto.dt.nona[is.na(auto.dt$horsepower)]$horsepower

fwrite(auto.dt.nona, 'C:/MSDS/MachL/Wk2/auto.dt.nona.csv')
```

Finally let's do some EDA on the data.  First a correlation plot, which shows all the variables to be highly correlated to mpg.
```{r corrplot}
corrplot(cor(auto.dt.nona))
```

There are not many 3- and 5-cyliner engines in the dataset, so we may want to throw those out.

```{r histograms}
dt.names <- names(auto.dt.nona)
for (i in seq(dim(auto.dt)[2])) {
  coldata <- auto.dt.nona[, get(dt.names[i])]
  n.levels <- nlevels(as.factor(coldata))
  if (n.levels <= 10) {
    barplot(table(coldata), xlab = dt.names[i])
  } else {
    hist(coldata, main = NULL, xlab = dt.names[i])
  }
  # add newlines so the plots all show up
  cat('\n\n')
}


```