---
title: "wk2 training"
author: "Jeanette Henry"
date: "January 22, 2018"
output: word_document
---

```{r setup}
library(data.table)
library(caret)
library(FNN)
library(class)
library(gmodels)

fn <- 'C:/MSDS/MachL/Wk2/heart.disease.data.clean.csv'

auto.dt <- fread(fn)
```

""	1. For the num column, leave the 0s as 0, and make anything that is 1 through 4 equal to 1.
	You can filter the rows of a data.table like this:
 heart.dt[num >=1, ]
to get the rows where num is greater than or equal to 1.
Then set the value of 'num' in those rows to 1.
I recommend using the := operator to set the num row by reference.""

```{r sort}
heart.dt<- fread(fn)
summary(heart.dt)
hist(heart.dt$num, breaks = 4)
heart.dt2 <- heart.dt
heart.dt2[num>=1, num:=1]
hist(heart.dt2$num)

```
"Split up the data into train/test datasets, I recommend using the createDataPartition() function from the caret library."
```{r split}
#1 set of data of targets to set apart from features
targs <- heart.dt2[, num]
# this selects all columns except mpg
features <- heart.dt2[, -'num', with=F]

set.seed(23)
trainIdx = createDataPartition(targs, p = 0.8)$Resample1
length(trainIdx)
length(targs)
tr.feats <- features[trainIdx]
tr.targs <- targs[trainIdx]
te.feats <- auto.dt[-trainIdx, -'num']
te.targs <- auto.dt[-trainIdx]$num

te.preds <- knn(tr.feats, te.feats, y = tr.targs, k = 3)
te.preds

library(gmodels)
CrossTable(x=iris.testLabels,y=iris_pred, prop.chisq = FALSE)

```

``` {r kvalues}
neighbors <- seq(2, 20)
r2s <- c()  # r-squared values (coefficient of determination)
sum.square.error <- c()  # sum of squares
for (k in neighbors) {
  preds <- knn.reg(train = tr.feats, y = tr.targs, k = k)
  r2s <- c(r2s, preds$R2Pred)
  sum.square.error <- c(sum.square.error, preds$PRESS)
}

# typically we look for an 'elbow' in the data, where the slope changes abruptly
# it's pretty clearly at 5 here
plot(neighbors, r2s)
plot(neighbors, sum.square.error)


```