---
title: "Week 2 inclass"
author: "Jeanette Henry"
date: "January 22, 2018"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(data.table)
library(DMwR)
library(corrplot)


fn <- 'C:/MSDS/MachL/Wk2/AirQualityUCI.csv'
aq<- fread(fn)
aqog<-fread(fn)

```



```{r aq}
str(aq)
aq = as.data.frame(sapply(aq, as.numeric))
str(aq)
```
NO2GT is now in numerics so we can use k nearest neighbor on it
``` {r knn}
summary(aq)
aq$V16 <- NULL
aq$V17<- NULL
aq$Date<- NULL
aq$Time<-NULL
aq$AH<-NULL
aq$RH<-NULL
aq$T<-NULL
aq$`C6H6(GT)`<-NULL


str(aq)
summary(aq)



hist(aq$`NO2(GT)`)
kaq<- knnImputation(aq)
```

## Including Plots

You can also embed plots, for example:

```{r original hist}
hist(aq$`NO2(GT)`)
```

```{R knn hist}
summary(kaq)
hist(kaq$`NO2(GT)`)
#this is the histogram with the k nearest neighbors completed

```

Now fill the NAs with the means, not using k nearest neighbors

```{r mean}
str(aq)
summary(aq)

library(zoo)
df2<-na.aggregate(aq)
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
###replace(df2, TRUE, lapply(df2, NA2mean))
df2[] <- lapply(df2, NA2mean)

summary(df2)

```
Now df2 is out data frame with the NA's replaced with the means

``` {r meanhist}

hist(df2$`NO2(GT)`)

```

plot

use original data table with the missing values to mask
use is.na to find the missing values

filter missing in original and grab that column and compare to where we imputted it 